\preface
\thispagestyle{empty}

Mathematics begins long before it becomes formal.

Before we learn to write an $\varepsilon$--$\delta$ proof, we already possess an intuitive sense of continuity. Before encountering groups and fields, we have learned to calculate, rearrange expressions, and recognize symmetry. Before topology gives a precise meaning to openness and connectedness, we can already distinguish, at least informally, between what is joined and what is separated.

Such intuition is not an obstacle to mathematics. It is where mathematics begins. Yet intuition alone is not enough. Familiar pictures may become misleading, plausible statements may fail in unexpected examples, and arguments that seem convincing may conceal assumptions that have never been stated. The development of mathematics is therefore marked by a persistent tension: we rely on intuition to discover what may be true, but we rely on definitions and proofs to determine what is true.

This manuscript grew out of my attempt to understand that passage from intuition to rigor.

It began as a collection of notes written while studying several foundational areas of undergraduate mathematics. At first, the purpose was practical: to record definitions, preserve useful arguments, and connect results that had appeared separately in lectures and textbooks. As the notes accumulated, however, a broader structure gradually emerged. Ideas from analysis, algebra, geometry, topology, and probability were not isolated constructions. They repeatedly relied on the same habits of thought: identifying the correct objects, choosing an appropriate notion of equivalence, determining which properties are preserved, and building general results from a small number of explicit assumptions.

The aim of this book is to make those connections visible.

It is not intended to be an encyclopedic treatment of mathematics, nor does it attempt to replace the many excellent textbooks devoted to each individual subject. Its scope is necessarily selective. Some theories are developed in considerable detail, while others appear only as introductions or as bridges to later study. The purpose is not to reach every advanced result, but to present a coherent route through several central mathematical languages and to explain why their basic definitions and theorems take the forms that they do.

The chapters therefore emphasize structure rather than isolated calculation. A definition is introduced not merely as terminology to be memorized, but as a response to a mathematical problem. A theorem is presented together with the mechanism that makes it work. Proofs are included not only to certify results, but also to reveal reusable ideas: approximation, compactness, completeness, decomposition, invariance, contradiction, and passage to a suitable limit. Examples illustrate the scope of a concept, while counterexamples mark the point beyond which an apparently natural statement ceases to be valid.

The organization of the book reflects a gradual movement toward abstraction. We begin with the language of logic and sets, since every later theory requires a precise way to speak about mathematical objects and their relations. From there, the discussion proceeds through analysis and linear algebra, where limits, continuity, vector spaces, and linear transformations provide some of the first substantial examples of general mathematical structure. Abstract algebra studies operations independently of the particular objects on which they act. Geometry and topology then examine space through increasingly flexible notions of distance, continuity, and invariance. Complex analysis reveals how an additional algebraic structure can produce unexpectedly strong analytic consequences. Measure theory and probability extend the language of size and integration to settings in which elementary methods are no longer sufficient.

These subjects differ in appearance, but they are held together by a common architectural principle. Mathematics advances by making its assumptions explicit, extracting their consequences through logical argument, and then asking which parts of the construction survive when the assumptions are weakened, altered, or transported to a new setting. In this sense, mathematical rigor is not the suppression of intuition. It is the discipline that allows intuition to travel safely beyond the examples from which it first arose.

The intended reader is a student who has encountered some university mathematics and wishes to see its separate courses as parts of a larger whole. The manuscript may be read consecutively, but it may also be used selectively. Earlier chapters provide language and techniques that are repeatedly invoked later, while each chapter is written to retain a degree of independence. Readers are encouraged to pause at definitions, reconstruct proofs before reading them, test hypotheses against examples, and ask what fails when a condition is removed. Mathematics is rarely learned by uninterrupted reading; it is learned by repeatedly attempting to reproduce an argument without the page.

No set of notes can eliminate every gap. Some proofs may admit shorter formulations, some explanations may be improved, and some choices of emphasis inevitably reflect the author's own path through the subject. This manuscript should therefore be regarded as a developing account rather than a final authority. Its purpose is fulfilled not when every line is accepted without question, but when the reader is led to examine the line more carefully, compare it with other sources, and continue the argument independently.

At the heart of mathematics remains the tension with which we began. We seek certainty, yet our questions continually lead us beyond what has already been formalized. We construct definitions to make thought precise, and those definitions in turn reveal possibilities that intuition alone would not have suggested. The mathematical edifice is built from axioms and proofs, but it is animated by curiosity.

This book is an invitation to take part in that construction: not merely to observe the finished theorems, but to understand why the foundations were laid as they were, how the different levels support one another, and where new structures might still be built.

\begin{flushright}
Jinshuo Li \\
Shanghai Jiao Tong University \\
2026.7.2, in Shanghai
\end{flushright}

\cleardoublepage
\pagenumbering{arabic}
\setcounter{page}{1}